---
title: "Learning Data Science"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE}
# collapsible heading at the top to link to other pages
```
<details>
  <summary>Posts</summary>

  [About Me](about_me.html)  
  
  [Why is Cable Media Obsessed with Joe Biden?](blog_post_one.html)
  
</details>  


As a data science student, [I](about_me.html) am constantly learning. With every class I take I am learning more technical skills, especially as I get into the nitty-gritty of coding. However, I am also in the process of learning soft skills along the way; being able to compute statistical analyses and create data visualizations is of little use without the knowledge of how to interpret this information. 


As a means to help me to continue strengthening my data analysis and data journalism skills, I have started this blog, where I will be posting articles discussing data analyses that I do in my free time as well as discussing what I am learning as I enter the field of data science.


Why should we care about data science? It is a tool that lets us talk about real issues that are happening all around us every day. Here, I explore some important things that I think we are not talking enough about. Explore some of my posts at the top to learn more. 
